The CrypTool Script: \\ Cryptography, Mathematics and more \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free eLearning program CrypTool\\ (8th edition -- distributed with CrypTool version 1.4.10)} \vspace{60pt}

### 2007-06-05 script/en/chapters/numbertheory.tex
Skript: Updates accord. to Vollmer's feedback about authors; C307 added (in German only yet)
Skript: updates, C307 Engl, mutual links)

\subsubsection*{Table 6: Values of $a^i{\rm ~mod~}46, 1\leq a,i<23$:}\label{SrcArith3c}
{ \textmd \small
\begin{center}  
\begin{tabular}{|p{16 pt}||@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|c|}
\hline
$a \setminus i$   & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 & 21 & 22 & 23 & ord \\
\hline
\hline
1    & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1 & 1    \\
\hline
2 & 2  & 4  & 8 & 16 & 32 & 18 & 36 & 26  & 6 & 12 & 24  & 2  & 4  & 8 & 16 & 32 & 18 & 36 & 26  & 6 & 12 & 24  & 2 & --    \\
\hline
3 & 3  & 9 & 27 & 35 & 13 & 39 & 25 & 29 & 41 & 31  & 1  & 3  & 9 & 27 & 35 & 13 & 39 & 25 & 29 & 41 & 31  & 1  & 3 & 11   \\
\hline
4  & 4 & 16 & 18 & 26 & 12  & 2  & 8 & 32 & 36  & 6 & 24  & 4 & 16 & 18 & 26 & 12  & 2  & 8 & 32 & 36  & 6 & 24  & 4 & --  \\
\hline
5 & 5 & 25 & 33 & 27 & 43 & 31 & 17 & 39 & 11  & 9 & 45 & 41 & 21 & 13 & 19  & 3 & 15 & 29  & 7 & 35 & 37  & 1  & 5 & 22  \\
\hline
6 & 6 & 36 & 32  & 8  & 2 & 12 & 26 & 18 & 16  & 4 & 24  & 6 & 36 & 32  & 8  & 2 & 12 & 26 & 18 & 16  & 4 & 24  & 6 & -- \\
\hline
7 & 7  & 3 & 21  & 9 & 17 & 27  & 5 & 35 & 15 & 13 & 45 & 39 & 43 & 25 & 37 & 29 & 19 & 41 & 11 & 31 & 33  & \textbf{1}  & 7 & 22 \\
\hline
8 & 8 & 18  & 6  & 2 & 16 & 36 & 12  & 4 & 32 & 26 & 24  & 8 & 18  & 6  & 2 & 16 & 36 & 12  & 4 & 32 & 26 & 24  & 8 & --  \\
\hline
9 & 9 & 35 & 39 & 29 & 31  & 3 & 27 & 13 & 25 & 41  & 1  & 9 & 35 & 39 & 29 & 31  & 3 & 27 & 13 & 25 & 41  & 1  & 9 & 11  \\
\hline
10 & 10  & 8 & 34 & 18 & 42  & 6 & 14  & 2 & 20 & 16 & 22 & 36 & 38 & 12 & 28  & 4 & 40 & 32 & 44 & 26 & 30 & 24 & 10 & --  \\
\hline 
11 & 11 & 29 & 43 & 13  & 5  & 9  & 7 & 31 & 19 & 25 & 45 & 35 & 17  & 3 & 33 & 41 & 37 & 39 & 15 & 27 & 21  & 1 & 11 & 22 \\
\hline
12 & 12  & 6 & 26 & 36 & 18 & 32 & 16  & 8  & 4  & 2 & 24 & 12  & 6 & 26 & 36 & 18 & 32 & 16  & 8  & 4  & 2 & 24 & 12 & -- \\
\hline
13 & 13 & 31 & 35 & 41 & 27 & 29  & 9 & 25  & 3 & 39  & 1 & 13 & 31 & 35 & 41 & 27 & 29  & 9 & 25  & 3 & 39  & 1 & 13 & 11  \\
\hline
14 & 14 & 12 & 30  & 6 & 38 & 26 & 42 & 36 & 44 & 18 & 22 & 32 & 34 & 16 & 40  & 8 & 20  & 4 & 10  & 2 & 28 & 24 & 14 & -- \\
\hline
15 & 15 & 41 & 17 & 25  & 7 & 13 & 11 & 27 & 37  & 3 & 45 & 31  & 5 & 29 & 21 & 39 & 33 & 35 & 19  & 9 & 43  & 1 & 15 & 22 \\
\hline
16 & 16 & 26  & 2 & 32  & 6  & 4 & 18 & 12  & 8 & 36 & 24 & 16 & 26  & 2 & 32  & 6  & 4 & 18 & 12  & 8 & 36 & 24 & 16 & -- \\
\hline
17 & 17 & 13 & 37 & 31 & 21 & 35 & 43 & 41  & 7 & 27 & 45 & 29 & 33  & 9 & 15 & 25 & 11  & 3  & 5 & 39 & 19  & 1 & 17 & 22 \\
\hline
18 & 18  & 2 & 36  & 4 & 26  & 8  & 6 & 16 & 12 & 32 & 24 & 18  & 2 & 36  & 4 & 26  & 8  & 6 & 16 & 12 & 32 & 24 & 18 & -- \\
\hline
19  & 19 & 39  & 5  & 3 & 11 & 25 & 15  & 9 & 33 & 29 & 45 & 27  & 7 & 41 & 43 & 35 & 21 & 31 & 37 & 13 & 17  & 1 & 19 & 22 \\
\hline
20  & 20 & 32 & 42 & 12 & 10 & 16 & 44  & 6 & 28  & 8 & 22 & 26 & 14  & 4 & 34 & 36 & 30  & 2 & 40 & 18 & 38 & 24 & 20 & --  \\
\hline
21 & 21 & 27 & 15 & 39 & 37 & 41 & 33  & 3 & 17 & 35 & 45 & 25 & 19 & 31  & 7  & 9  & 5 & 13 & 43 & 29 & 11  & 1 & 21 & 22  \\
\hline
22 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & -- \\
\hline
23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & -- \\
\hline 
\end{tabular}
\end{center}
}

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\hypertarget{Chapter_ElementaryNT_10}{}
\subsection{Proof of the RSA procedure with Euler-Fermat}
\index{RSA}\label{rsabeweis}

\hypertarget{RSABeweis} Using the Euler-Fermat theorem, we can ``prove'' the
RSA\footnote{The RSA procedure\index{RSA!RSA procedure} is the most
common asymmetric cryptography procedure. Developed in 1978 by Ronald Rivest,
Adi Shamir and Leonard Adleman, it can be used both for signatures and for
encryption. Cryptographers always associate this procedure with the
abbreviation ``RSA'' $-$ the following remark is meant with humor to show
that each letter combination can be used with several meanings: In Britain
the ``Royal Society for the encouragement of Arts, Manufactures \& Commerce''
is commonly known as the ``RSA''.
}$^,$\footnote{%
In literature\index{Literature} and in movies\index{Movies} not only 
classic but also modern cryptographic methods have been used 
%(\hyperlink{appendix-movies}
%{see appendix A3: Movies and Literature with Relation to Cryptography}).
(see appendix \ref{s:appendix-movies}).
% hier die items nicht einrücken ! xxxxxxxxxxxxxxx
% \begin{list}{\textbullet}{\leftmargin10pt}
%   \item the movie ``House of Cards'', 1992
%   \item the story ``The Dialogue of the Sisters'' by Dr.~C.~Elsner
% \end{list}\vspace{-\baselineskip}\vspace{-\parskip}
} %end footnote
procedure in the group $\mathbb{Z}_n^*$.

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection[Considerations regarding the security of the RSA algorithm]
{Considerations regarding the security of the RSA algorithm\footnotemark}
    \footnotetext{Major
    parts of chapters \ref{complexity} and \ref{chptSecurityParam} follow
    the article ``Vorz\"uge und Grenzen des RSA-Verfahrens'' written by
    F.~Bourseau, D.~Fox and C.~Thiel \cite{Bourseau2002}
             }
    \label{SecurityRSA}

@@

The web pages
\href{http://www.crypto-world.com}{\texttt{http://www.crypto-world.com}}\ 
and
\href{http://www.tutorgig.com/ed/RSA\_number}{\texttt{http://www.tutorgig.com/ed/RSA\_number}}\footnote{%
This site was not quite up-to-date end of May 2005: RSA-200 was missing.}
contain exhaustive overviews about the factoring 
records\index{Factorisation!factoring records} of composed
integers using different methods.

\vskip +10pt
The current record (as of May 2005) obtained using the GNFS method
(General Number Field Sieve) factorised a 200 decimal digit into its
both prime factors.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|c|ccccc|}
\hline
	& {\bf Decimal digits} & {\bf Binary digits} & & {\bf Factored on} & {\bf Factored by} \\
\hline
	RSA-155	&	155 & 512 & & August, 1999 & Herman te Riele et al. \\
	C158 & 		158 & 523 & & January, 2002 & Jens Franke et al. \\
	RSA-160 & 	160 & 530 & & April, 2003 & Jens Franke et al. \\
	RSA576 & 	174 & 576 & & December, 2003 & Jens Franke et al. \\
	C176 & 		176 & 583 & & May, 2005 & Kazumaro Aoki et al. \\
	RSA640\footnotemark & 	193 & 600 & & November, 2005 & Jens Franke et al. \\
	RSA-200 & 	200 & 663 & & May, 2005 & Jens Franke et al. \\
\hline
\end{tabular}
\caption{The current general purpose factoring records (as of November 2005)}    % Eyecatcher_neue-Mersenne 
\end{center}
\end{table} 
\footnotetext{%
A research group of the GISA solved this challenge which was awarded with
20,000 US dollar using the GNFS method. The researchers needed about five
months to divide this number into its both 320 bit long prime factors.\\
RSA Labs offers its challenges since the beginning of the 1990th.
The next bigger challenge awarded with 30,000 US dollars is RSA-704.\\
See:
\href{http://www.heise.de/newsticker/meldung/print/65957}
{\texttt{http://www.heise.de/newsticker/meldung/print/65957}}
}

@@

Below the last records are explained in more detail.

@@

The last remaining factor, called ``C158'', was known to be composite by
then, but its factors where not known (the following 3 lines contain one number):
$$
\begin{array}{c}
39505874583265144526419767800614481996020776460304936 \\
45413937605157935562652945068360972784246821953509354 \\
4305870490251995655335710209799226484977949442955603
\end{array}
$$
The factorisation of C158 resulted in the following two 73- and 86-digit prime factors:
$$
\begin{array}{c}
3388495837466721394368393204672181522815830368604993048084925840555281177
\end{array}
$$
and
$$
\begin{array}{c}
1165882340667125990314837655838327081813101 \\
2258146392600439520994131344334162924536139.
\end{array}
$$
So now all 8 prime factors of $2^{953} - 1$ have been found.

\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item[]   \href{http://www.loria.fr/~zimmerma/records/gnfs158}
       {\texttt{http://www.loria.fr/~zimmerma/records/gnfs158}} \\
          \href{http://www.crypto-world.com/FactorRecords.html}
       {\texttt{http://www.crypto-world.com/FactorRecords.html}} \\
          \href{http://www.crypto-world.com/announcements/c158.txt}
       {\texttt{http://www.crypto-world.com/announcements/c158.txt}}
\end{itemize}
\end{minipage}
\vspace{12pt}

@@

The calculations took place between December 2003 and May 2005.
The factorization done by the group around Bahr, B\"ohm, Franke, Kleinjung, 
Montgomery and te Riele lasted almost 17 months.
The operating expense of the calculations was about 120,000 
MIPS-years\footnote{%
A MIPS-year (MY) is the quantity of operations a machine can perform in one year,
if the machine constantly achieves one million integer operations per second (MIPS).
For illustration: a INTEL Pentium 100 processor achieves about 50 MIPS.
To factorize a 2048 bit module it is estimated to need about {$8.5 \cdot
 10^{40}$ MY}. }.

% --------------------------------------------------------------------------
\vskip +50pt
\paragraph{Size of factorized numbers compared to primality proven numbers}
\mbox{}

@@

% --------------------------------------------------------------------------
\vskip +60pt
\subsubsection{Further current research about primes and factorisation}
Prime numbers are part of very many topical research areas in number theory
and computer science. Progress made with factorisation is bigger than was
estimated 5 years ago -- this is not only due to faster computers but
also new knowledge.

@@

\bibitem[Buchmann1999]{Buchmann1999} \index{Buchmann 1999}
    Johannes Buchmann, \\
    {\em Einf\"uhrung in die Kryptographie}, Springer, 1999.

@@

\bibitem[GISA2002]{GISA2002} \index{GISA 2002}
    GISA (German Information Security Agency), \\
    {\em Recommendation for key length selection}, Bonn, Sep. 2002, \\
    \href{http://www.bsi.bund.de/esig/basics/techbas/krypto/bund02v7.pdf}
    {\tt http://www.bsi.bund.de/esig/basics/techbas/krypto/bund02v7.pdf} \\
    A statement on these recommendations: \\
    \href{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}
    {\texttt{http://www.secorvo.de/publikat/}}\\
    \vspace*{1cm}
    \href{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}
    {\texttt{stellungnahme-algorithmenempfehlung-020307.pdf}}

>>>>

\subsubsection*{Table 6: Values of $a^i{\rm ~mod~}46, 1\leq a,i<23$:}
\label{SrcArith3c}
\textmd \small
\begin{center}  
\begin{tabular}{|p{16 pt}||@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|@{\:}r@{\:}|c|}
\hline
$a \setminus i$   & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 & 21 & 22 & 23 & ord \\
\hline
\hline
1    & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1  & 1 & 1    \\
\hline
2 & 2  & 4  & 8 & 16 & 32 & 18 & 36 & 26  & 6 & 12 & 24  & 2  & 4  & 8 & 16 & 32 & 18 & 36 & 26  & 6 & 12 & 24  & 2 & --    \\
\hline
3 & 3  & 9 & 27 & 35 & 13 & 39 & 25 & 29 & 41 & 31  & 1  & 3  & 9 & 27 & 35 & 13 & 39 & 25 & 29 & 41 & 31  & 1  & 3 & 11   \\
\hline
4  & 4 & 16 & 18 & 26 & 12  & 2  & 8 & 32 & 36  & 6 & 24  & 4 & 16 & 18 & 26 & 12  & 2  & 8 & 32 & 36  & 6 & 24  & 4 & --  \\
\hline
5 & 5 & 25 & 33 & 27 & 43 & 31 & 17 & 39 & 11  & 9 & 45 & 41 & 21 & 13 & 19  & 3 & 15 & 29  & 7 & 35 & 37  & 1  & 5 & 22  \\
\hline
6 & 6 & 36 & 32  & 8  & 2 & 12 & 26 & 18 & 16  & 4 & 24  & 6 & 36 & 32  & 8  & 2 & 12 & 26 & 18 & 16  & 4 & 24  & 6 & -- \\
\hline
7 & 7  & 3 & 21  & 9 & 17 & 27  & 5 & 35 & 15 & 13 & 45 & 39 & 43 & 25 & 37 & 29 & 19 & 41 & 11 & 31 & 33  & \textbf{1}  & 7 & 22 \\
\hline
8 & 8 & 18  & 6  & 2 & 16 & 36 & 12  & 4 & 32 & 26 & 24  & 8 & 18  & 6  & 2 & 16 & 36 & 12  & 4 & 32 & 26 & 24  & 8 & --  \\
\hline
9 & 9 & 35 & 39 & 29 & 31  & 3 & 27 & 13 & 25 & 41  & 1  & 9 & 35 & 39 & 29 & 31  & 3 & 27 & 13 & 25 & 41  & 1  & 9 & 11  \\
\hline
10 & 10  & 8 & 34 & 18 & 42  & 6 & 14  & 2 & 20 & 16 & 22 & 36 & 38 & 12 & 28  & 4 & 40 & 32 & 44 & 26 & 30 & 24 & 10 & --  \\
\hline 
11 & 11 & 29 & 43 & 13  & 5  & 9  & 7 & 31 & 19 & 25 & 45 & 35 & 17  & 3 & 33 & 41 & 37 & 39 & 15 & 27 & 21  & 1 & 11 & 22 \\
\hline
12 & 12  & 6 & 26 & 36 & 18 & 32 & 16  & 8  & 4  & 2 & 24 & 12  & 6 & 26 & 36 & 18 & 32 & 16  & 8  & 4  & 2 & 24 & 12 & -- \\
\hline
13 & 13 & 31 & 35 & 41 & 27 & 29  & 9 & 25  & 3 & 39  & 1 & 13 & 31 & 35 & 41 & 27 & 29  & 9 & 25  & 3 & 39  & 1 & 13 & 11  \\
\hline
14 & 14 & 12 & 30  & 6 & 38 & 26 & 42 & 36 & 44 & 18 & 22 & 32 & 34 & 16 & 40  & 8 & 20  & 4 & 10  & 2 & 28 & 24 & 14 & -- \\
\hline
15 & 15 & 41 & 17 & 25  & 7 & 13 & 11 & 27 & 37  & 3 & 45 & 31  & 5 & 29 & 21 & 39 & 33 & 35 & 19  & 9 & 43  & 1 & 15 & 22 \\
\hline
16 & 16 & 26  & 2 & 32  & 6  & 4 & 18 & 12  & 8 & 36 & 24 & 16 & 26  & 2 & 32  & 6  & 4 & 18 & 12  & 8 & 36 & 24 & 16 & -- \\
\hline
17 & 17 & 13 & 37 & 31 & 21 & 35 & 43 & 41  & 7 & 27 & 45 & 29 & 33  & 9 & 15 & 25 & 11  & 3  & 5 & 39 & 19  & 1 & 17 & 22 \\
\hline
18 & 18  & 2 & 36  & 4 & 26  & 8  & 6 & 16 & 12 & 32 & 24 & 18  & 2 & 36  & 4 & 26  & 8  & 6 & 16 & 12 & 32 & 24 & 18 & -- \\
\hline
19  & 19 & 39  & 5  & 3 & 11 & 25 & 15  & 9 & 33 & 29 & 45 & 27  & 7 & 41 & 43 & 35 & 21 & 31 & 37 & 13 & 17  & 1 & 19 & 22 \\
\hline
20  & 20 & 32 & 42 & 12 & 10 & 16 & 44  & 6 & 28  & 8 & 22 & 26 & 14  & 4 & 34 & 36 & 30  & 2 & 40 & 18 & 38 & 24 & 20 & --  \\
\hline
21 & 21 & 27 & 15 & 39 & 37 & 41 & 33  & 3 & 17 & 35 & 45 & 25 & 19 & 31  & 7  & 9  & 5 & 13 & 43 & 29 & 11  & 1 & 21 & 22  \\
\hline
22 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & 24 & 22 & -- \\
\hline
23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & 23 & -- \\
\hline 
\end{tabular}
\end{center}
}

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\hypertarget{Chapter_ElementaryNT_10}{}
\hypertarget{RSABeweis}{}
\subsection{Proof of the RSA procedure with Euler-Fermat}
\index{RSA}\label{rsabeweis}
Using the Euler-Fermat theorem, we can ``prove'' the
RSA\footnote{The RSA procedure\index{RSA!RSA procedure} is the most
common asymmetric cryptography procedure. Developed in 1978 by Ronald Rivest,
Adi Shamir and Leonard Adleman, it can be used both for signatures and for
encryption. Cryptographers always associate this procedure with the
abbreviation ``RSA'' $-$ the following remark is meant with humor to show
that each letter combination can be used with several meanings: In Britain
the ``Royal Society for the encouragement of Arts, Manufactures \& Commerce''
is commonly known as the ``RSA''.
}$^,$\footnote{%
In literature\index{Literature} and in movies\index{Movies} not only 
classic but also modern cryptographic methods have been used 
%(\hyperlink{appendix-movies}
%{see appendix A3: Movies and Literature with Relation to Cryptography}).
(see appendix \ref{s:appendix-movies}).
% hier die items nicht einrücken ! xxxxxxxxxxxxxxx
% \begin{list}{\textbullet}{\leftmargin10pt}
%   \item the movie ``House of Cards'', 1992
%   \item the story ``The Dialogue of the Sisters'' by Dr.~C.~Elsner
% \end{list}\vspace{-\baselineskip}\vspace{-\parskip}
} %end footnote
procedure in the group $\mathbb{Z}_n^*$.

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\hypertarget{SecurityRSA}{}
\subsection[Considerations regarding the security of the RSA algorithm]
{Considerations regarding the security of the RSA algorithm\footnotemark}
    \footnotetext{Major
    parts of chapters \ref{complexity} and \ref{chptSecurityParam} follow
    the article ``Vorz\"uge und Grenzen des RSA-Verfahrens'' written by
    F.~Bourseau, D.~Fox and C.~Thiel \cite{Bourseau2002}.}
    \label{SecurityRSA}

@@

An exhaustive overview about the factoring records
\index{Factorisation!factoring records} of composed integers using different
methods can be found on the following web pages:
\vspace{-10pt}
\begin{itemize}
\item[]
     \href{http://www.crypto-world.com}
  {\texttt{http://www.crypto-world.com}} \\
     \href{http://www.tutorgig.com/ed/RSA\_number}
  {\texttt{http://www.tutorgig.com/ed/RSA\_number}}  ~~ The RSA Factoring Challenge \\
     \href{http://en.wikipedia.org/wiki/Integer_factorization_records}
  {\texttt{http://en.wikipedia.org/wiki/Integer\_factorization\_records}}
\end{itemize}
%\footnote{%
%This site was not quite up-to-date end of May 2005: RSA-200 was missing.}

The current record (as of May 2005) obtained using the GNFS method
(General Number Field Sieve) factorised a general 200 decimal digit into its
both prime factors.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|c|ccccc|}
\hline
	& {\bf Decimal digits} & {\bf Binary digits} & & {\bf Factored on} & {\bf Factored by} \\
\hline
	RSA-155	&	155 & 512 & & August, 1999 & Herman te Riele et al. \\
	C158 & 		158 & 523 & & January, 2002 & Jens Franke et al. \\
	RSA-160 & 	160 & 530 & & April, 2003 & Jens Franke et al. \\
	RSA576 & 	174 & 576 & & December, 2003 & Jens Franke et al. \\
	C176 & 		176 & 583 & & May, 2005 & Kazumaro Aoki et al. \\
	RSA640\footnotemark & 	193 & 600 & & November, 2005 & Jens Franke et al. \\
	RSA-200 & 	200 & 663 & & May, 2005 & Jens Franke et al. \\
	C307 & 		307 & 1017 & & May, 2007 & Jens Franke et al. \\
\hline
\end{tabular}
\caption{The current factoring records (as of June 2007)}    % Eyecatcher_neue-Mersenne 
\end{center}
\end{table} 
\footnotetext{%
A research group of the GISA solved this challenge which was awarded with
20,000 US dollar using the GNFS method. The researchers needed about five
months to divide this number into its both 320 bit long prime factors.\\
RSA Labs offers its challenges since the beginning of the 1990th.
The next bigger challenge awarded with 30,000 US dollars is RSA-704.\\
See:
\href{http://www.heise.de/newsticker/meldung/print/65957}
{\texttt{http://www.heise.de/newsticker/meldung/print/65957}}
}

@@

Below the last records are explained in more detail.
The two methods, GNFS and SNFS, used to do so are shortly illustrated at the 
following web pages:
\index{General Number Field Sieve (GNFS)}
\index{Special Number Field Sieve (SNFS)}
\vspace{-10pt}
\begin{itemize}
\item[]
      \href{http://en.wikipedia.org/wiki/Special_number_field_sieve}
   {\texttt{http://en.wikipedia.org/wiki/Special\_number\_field\_sieve}} \\
      \href{http://en.wikipedia.org/wiki/General_number_field_sieve}
   {\texttt{http://en.wikipedia.org/wiki/General\_number\_field\_sieve}}
\end{itemize}
%\vspace{-10pt}

@@

The last remaining factor, called ``C158'', was known to be composite by
then, but its factors were not known (the following 3 lines contain one number):
$$
\begin{array}{c}
39505874583265144526419767800614481996020776460304936 \\
45413937605157935562652945068360972784246821953509354 \\
4305870490251995655335710209799226484977949442955603
\end{array}
$$
The factorisation of C158 resulted in the following two 73- and 86-digit prime factors:
$$
\begin{array}{c}
3388495837466721394368393204672181522815830368604993048084925840555281177
\end{array}
$$
and
$$
\begin{array}{c}
1165882340667125990314837655838327081813101 \\
2258146392600439520994131344334162924536139.
\end{array}
$$
So now all 8 prime factors of $2^{953} - 1$ have been found.

\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item[]   \href{http://www.loria.fr/~{}zimmerma/records/gnfs158}
       {\texttt{http://www.loria.fr/~{}zimmerma/records/gnfs158}} \\
          \href{http://www.crypto-world.com/FactorRecords.html}
       {\texttt{http://www.crypto-world.com/FactorRecords.html}} \\
          \href{http://www.crypto-world.com/announcements/c158.txt}
       {\texttt{http://www.crypto-world.com/announcements/c158.txt}}
\end{itemize}
\end{minipage}
\vspace{12pt}

@@

The calculations took place between December 2003 and May 2005.
The factorization done by the group around Bahr, B\"ohm, Franke, Kleinjung, 
Montgomery and te Riele lasted almost 17 months.
The operating expense of the calculations was about 120,000 
MIPS-years\footnote{%
A MIPS-year (MY) is the quantity of operations a machine can perform in one year,
if the machine constantly achieves one million integer operations per second (MIPS).
For illustration: a INTEL Pentium 100 processor achieves about 50 MIPS.
To factorize a 2048 bit module it is estimated to need about {$8.5 \cdot
 10^{40}$ MY}. }.
\vspace{24pt}

% --------------------------------------------------------------------------
\vskip +20pt
\paragraph{C307 / M1039} \label{C307} \index{C307} \index{M1039}\mbox{}
\hypertarget{C307-chap3}{}

In May 2007 Prof. Franke, Prof. Kleinjung (University of Bonn),
the Japanese telecommunication company NTT and Prof. Arjen Lenstra
(Polytechnical University of Lausanne) announced, that they managed to
factorize a $307$ digit decimal number into its both prime factors
with the SNFS method (Special Number Field Sieve)
\index{Special Number Field Sieve (SNFS)} within 11 months
(the two factors have 80 and 227 decimal digits).

The task of the researchers was not initiated by a challenge, but
they wanted to find the last prime factors of the Mersenne number $2^{1039}+1$
(see ``Wanted List'' of the Cunningham Project\index{%
Cunningham project}\footnote{%
Cunningham project: \href{http://www.cerias.purdue.edu/homes/ssw/cun/}
                 {\texttt{http://www.cerias.purdue.edu/homes/ssw/cun/}}\\
Cunningham table: \href{http://homes.cerias.purdue.edu/~ssw/cun/pmain1206}
                {\texttt{http://homes.cerias.purdue.edu/\~{}ssw/cun/pmain1206}}\\
The numbers in the Cunningham table have the following syntax:\\
``(2,n)-'' means $2^{n}-1$;~~~
``(2,n)+'' means $2^{n}+1$.\\
To describe the magnitude one writes $p<n>$ or $c<n>$: 
``n'' is the number of decimal digits and ``p'' and ``c'' tell,
whether the number is prime or composite.\\
$2^{1039}-1 = p7 * c307 = p7 * p80 * p227$ \\
It is explained more precisely at the CUN page:\\
``2,651+ means $2^{651} + 1$ and the size (c209 means 209 decimal digits)
of the
number which was factored.  Then come the new factor(s), the discoverer and
the method used.  Recently, only the multiple polynomial quadratic sieve
(ppmpqs), the elliptic curve method (ecm) and the number field sieve (nfs)
have been used.  `hmpqs' stands for hypercube multiple polynomial quadratic
sieve.  Under `new factors', `p90' means a 90-digit prime and `c201' is a
201-digit composite number.''.
}).\\

The number $2^{1039}-1$ consists of 3 prime factors: The smallest one, 
$p7 = 5080711$ was already known.\footnote{%
This one can also be found using CrypTool\index{CrypTool}  via menu 
{\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem \textbackslash{} 
Factorisation of a Number} --- with the algorithms of Brent, Williams or
Lenstra, which are ``relativ'' good to separate small factors. \\
CrypTool can factorize in a reasonable time numbers no longer than 250 bit.
}

To complete this the second factor (co-divider) ``C307'' had to be factorized:
Till then it was only known, that the last remaining factor was composite,
but it was unknown, how many prime factors it had and what are the prime factors.
The following 5 lines contain one number:
$$
\begin{array}{c}
C307 =1159420574072573064369807148876894640753899791702017724986868353538\\
8224838599667566080006095408005179472053993261230204874402860435302\\
8619141014409345351233471273967988850226307575280937916602855510550\\
0425810771176177610094137970787973806187008437777186828680889844712\\
822002935201806074755451541370711023817
\end{array}
$$
The factorisation of C307 resulted in the following two 80- and 2276-digit prime factors:
$$
\begin{array}{c}
p80 = 558536666199362912607492046583159449686465270184\\
      88637648010052346319853288374753
\end{array}
$$
and
$$
\begin{array}{c}
p227 = 207581819464423827645704813703594695162939708007395209881208\\
       387037927290903246793823431438841448348825340533447691122230\\
       281583276965253760914101891052419938993341097116243589620659\\
       72167481161749004803659735573409253205425523689
.
\end{array}
$$
So now the number $2^{1039}-1$ is completely factorized in its 3 prime factors.

\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item[]   \href{http://www.loria.fr/~zimmerma/records/21039-}
       {\texttt{http://www.loria.fr/\~{}zimmerma/records/21039-}} \\
          \href{http://www.crypto-world.com/announcements/m1039.txt}
       {\texttt{http://www.crypto-world.com/announcements/m1039.txt}} \\
          \href{http://www.crypto-world.com/FactorAnnouncements.html}
       {\texttt{http://www.crypto-world.com/FactorAnnouncements.html}} \\
          \href{http://www1.uni-bonn.de/pressDB/jsp/pressemitteilungsdetails.jsp?detailjahr=2007&detail=160}
       {\texttt{http://www1.uni-bonn.de/pressDB/jsp/pressemitteilungsdetails.jsp?}} \\
          \href{http://www1.uni-bonn.de/pressDB/jsp/pressemitteilungsdetails.jsp?detailjahr=2007&detail=160}
       {\texttt{\quad \quad \quad detailjahr=2007\&detail=160}}\end{itemize}
\end{minipage}

% --------------------------------------------------------------------------
\vskip +70pt
\paragraph{Size of factorized numbers compared to primality proven numbers}
\mbox{}

@@

% --------------------------------------------------------------------------
\vskip +60pt
\subsubsection{Further current research about primes and factorisation}
\label{FactorisationResearch}
Prime numbers are part of very many topical research areas in number theory
and computer science. Progress made with factorisation is bigger than was
estimated 5 years ago -- this is not only due to faster computers but
also new knowledge.

@@

\bibitem[Buchmann2004]{Buchmann2004} \index{Buchmann 2004}
    Johannes Buchmann, \\
    {\em Introduction to Cryptography}, Springer, 2nd edition, 2004.

@@

\bibitem[GISA2002]{GISA2002} \index{GISA 2002}
    GISA (German Information Security Agency), \\
    {\em Recommendation for key length selection}, Bonn, Sep. 2002, \\
    \href{http://www.bsi.bund.de/esig/basics/techbas/krypto/bund02v7.pdf}
    {\tt http://www.bsi.bund.de/esig/basics/techbas/krypto/bund02v7.pdf} \\
    A statement on these recommendations: \\
    \href{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}
    {\texttt{http://www.secorvo.de/publikat/}}\\
    \hspace*{1cm}
    \href{http://www.secorvo.de/publikat/stellungnahme-algorithmenempfehlung-020307.pdf}
    {\texttt{stellungnahme-algorithmenempfehlung-020307.pdf}}
